---
title: "SIDER - The Side Effect Resource"
author: "Martin Frigaard"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
library(remedy)
# figs folder
if (!file.exists("figs")) {
  dir.create("figs")
}
# chunk options
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/", # location of files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
) # width of figures
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
```

# Motivation

This document outlines how we found side effects, adverse events, clinical trials, and other information for the following covid-19 treatments: Ivermectin, Remdesivir, Tocilizumab, and Hydroxychloroquine + Azithromycin

## SIDER database

These data come from the SIDER database. See links below:

- The database: http://sideeffects.embl.de/

- Journal article: http://nar.oxfordjournals.org/content/early/2015/10/19/nar.gkv1075.abstract

- Example of [Fluoxetine](http://sideeffects.embl.de/drugs/3386/).

## Import the data

```{r import, message=FALSE, warning=FALSE}
DrugNames <- readr::read_delim(file = "data/sider-data/drug_names.tsv",
                               delim = "\t",
                               escape_double = FALSE,
                               col_names = FALSE,
                               trim_ws = FALSE) %>%
  # set names
             purrr::set_names(x = .,
                              nm = "drug_cid", "drug_name")

DrugAtc <- read_delim(file = "data/sider-data/drug_atc.tsv",
                      delim = "\t",
                      escape_double = FALSE,
                      col_names = FALSE,
                      trim_ws = FALSE) %>%
            purrr::set_names(x = .,
                              nm = "drug_cid", "drug_atc")
```

Join these two tables on CID

```{r DrugCIDATC}
DrugCIDATC <- DrugAtc %>%
  left_join(x = ., y = DrugNames, by = "drug_cid")
DrugCIDATC %>% glimpse()
```

Import the `MEDDRA` side effect data

```{r MeddraAllSe, message=FALSE, warning=FALSE}
MeddraAllSe <- readr::read_delim("data/sider-data/meddra_all_se.tsv", delim = "\t",
                                 escape_double = FALSE,
                                 col_names = FALSE,
                                 trim_ws = TRUE) %>%
              purrr::set_names(x = .,
                            nm = "STITCH_compound_id01", "STITCH_compound_id02",
                            "UMLS_concept_id", "MedDRA_concept_type",
                            "UMLS_concept_id_for_MedDRA_term",
                            "side_effect_name")

head(MeddraAllSe)
```

Now we can see which drugs have side effects in the SIDER database.

### Ivermectin

First we look up the ID in the 

```{r IvermectinID}
IvermectinID <- DrugCIDATC %>% 
  dplyr::filter(drug_atc == "P02CF01")
IvermectinID
```


```{r IvermectinID-STITCH_compound_id01-STITCH_compound_id02}
IvermectinAllSe <- MeddraAllSe %>% 
  dplyr::filter(STITCH_compound_id01 == "CID106435110" | 
                  STITCH_compound_id02 == "CID106435110")
utils::head(IvermectinAllSe)
```


### Remdesivir

This drug does not have an ATC number, but Wikipedia lists a PubChem CID of `121304016`

```{r remdesivir-check-cid}
MeddraAllSe %>% 
  dplyr::filter(STITCH_compound_id01 == "121304016")
```


### Tocilizumab

### Hydroxychloroquine   

### Azithromycin

