---
title: "Ojo publico - drug side effects"
author: "Martin Frigaard"
date: "`r Sys.Date()`"
output: 
  github_document: 
    df_print: kable
always_allow_html: true
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
library(hrbrthemes)
library(remedy)
# figs folder
if (!file.exists("figs")) {
  dir.create("figs")
}
# chunk options
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/", # location of files
  fig.height = 7.5, # height of figures
  fig.width = 10 # width of figures
) # width of figures
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
```

# Motivation

This document outlines how we found side effects, adverse events, clinical trials, and other information for the following covid-19 treatments: Ivermectin, Remdesivir, Tocilizumab, and Hydroxychloroquine + Azithromycin

## Medlineplus Googlesheet data

The code below imports data from a Google Drive spreadsheet of drugs and side effects collected from [medlineplus.gov](https://medlineplus.gov/druginformation.html) and [medscape](https://reference.medscape.com/).

```{r side-effects}
# devtools::install_github("tidyverse/googlesheets4")
library(googlesheets4)
gs4_deauth()
```

```{r SideEffects}
SideEffects <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/12_4fGi2UNnnMiWvhn8VNTGi0mXhl5bf_h1Swo8xG2Sc/edit#gid=0", 
                                         sheet = "side-effects") %>% janitor::clean_names()
SideEffects <- SideEffects %>% dplyr::glimpse(78)
```

Start by wrangling the `Azithromycin` side effects. 

```{r AzithromycinSE}
AzithromycinSE <- SideEffects %>% 
  filter(drug == "Azithromycin") %>% 
  select(drug, medscape_adverse_effects)
writeLines(AzithromycinSE$medscape_adverse_effects)
```

I want to split this on the semi-colon `;` and new line `\n`, which I can identify with the pattern below:

```{r test-AzithromycinSE-pattern-01}
str_view_all(AzithromycinSE$medscape_adverse_effects, pattern = "\\; ")
```


```{r}
AzithromycinSE %>%  
        dplyr::mutate(medscape_adverse_effects = 
                        # 1) split this on the "updated" pattern
             stringr::str_split(string = medscape_adverse_effects, 
                                pattern = "\\; ")) %>% 
                       # 2) convert the output from split into multiple rows
        tidyr::unnest(medscape_adverse_effects) %>% 
    # check this to see what is happening 
    dplyr::select(drug, medscape_adverse_effects) %>% 
    # remove \n
     dplyr::mutate(medscape_adverse_effects = 
                     stringr::str_remove_all(string = medscape_adverse_effects, 
                                             pattern = "\n"))
```


## Datawrapper table format

This format will include "Drug, Class, Adverse effects, Comorbidities"

```{r test-x}
# adverse events
CovidDrugAdverseEvents <- CovidDrugs %>%
  dplyr::select(Drug = drug, 
                side_effect) %>% 
  dplyr::group_by(Drug) %>%
  dplyr::summarize(
    `Adverse effects` = paste(side_effect, collapse = ', ')) %>% 
  dplyr::mutate(`Adverse effects` = stringr::str_to_sentence(`Adverse effects`))
CovidDrugAdverseEvents
# Comorbidities
CovidDrugComorbidities <- CovidDrugs %>% 
  dplyr::select(Drug = drug, 
                primary_comorbidity, 
                secondary_comorbidity) %>% 
  tidyr::unite(primary_comorbidity, secondary_comorbidity, 
               col = "Comorbidities", sep = ", ") %>% 
  dplyr::mutate(Comorbidities = stringr::str_to_sentence(Comorbidities)) %>%
  dplyr::distinct() 
CovidDrugComorbidities
# Drug and class
CovidDrugClass <- CovidDrugs %>% 
  dplyr::select(Drug = drug, 
                Class = drug_class) %>% 
  dplyr::mutate(Class = stringr::str_to_sentence(Class)) %>%
  dplyr::distinct()
# join
DataWrapperCovidDrugs <- dplyr::inner_join(CovidDrugClass, 
           CovidDrugAdverseEvents, by = "Drug") %>% 
  dplyr::inner_join(x = ., y = CovidDrugComorbidities, by = "Drug")
knitr::kable(DataWrapperCovidDrugs)
```

Export these data as .csv

```{r export-DataWrapperCovidDrugs}
fs::dir_create("data/medline-side-effects")
readr::write_csv(as.data.frame(DataWrapperCovidDrugs), 
                 path = "data/medline-side-effects/DataWrapperCovidDrugs.csv")
```

